### 2021-02-01 abstract/IQFA 2020/fiche.tex

\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usetikzlibrary{optics}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}
\usepackage[margin=0in]{geometry}

\title{Titre}
\author{Clément Pellet-Mary}
\date\today

\begin{document}
\centering
\textbf{Abstract :} The nitrogen-vacancy center's spin is a powerful sensor used in a large variety of applications. Here we use Cross-Relaxations (C.R.) to detect the presence of dark spins (VH$^-$ and WAR1) in CVD-grown diamond. We also managed to mechanically detect cross-relaxations between two classes of NV centers by measuring the torque applied on a levitating diamond.
  \end{document}