### 2021-04-13 Mes papiers/CR-CVD/eproof_answer.tex
tutu

\documentclass{article}
\usepackage{physics}
\usepackage[]{quoting}

\begin{document}

\begin{enumerate}
\item   Please clarify what “it” refers to in “is that it can be optically.” NV-?

"it" refers to the NV- spin
\item   Please clarify what “it” refers to in “It would, indeed, offer the.” “This technology”?

"it" refers to this technology
\item   Please define VH-and WAR1

VH- corresponds to the negatively charged vacancy-hydrogen center

WAR1 corresponds to the first unidentified diamond spin defect found at the University of Warwick
\item   Please clarify what “this” refers to in “This could otherwise result in.” “This projection”?

"this" refers to "this method"
\item   The asterisk was deleted from Table I because it did not seem to correspond to anything. If it had a mean, Please define.and check formating of the table

the asterisk had no meaning
\item   Correct to use 5 rather than V for the section of the SM to match the other references. Please check.

It is correct to use 5 rather than V (note pour Gabriel : en vrai le SI est écrit en chiffre romain, est-ce que ça vaut le coup de le signaler (et de changer toutes les refs sauf celle-ci))
\item   Please define NVH

NVH stands for the nitrogen-vacancy-hydrogen center
\item   Please define ZFS

ZFS stands for zero-field splitting
\item   Please provide page/citation number for Ref. [4]

The page numbers are 131-156.
\item   Please update Ref. [15] with a published article if possible

ref. [15] has not yet been published in a peer-reviewed journal
\item   Please provide a description of the Supplemental Material in Ref. [21]

"for further experimental and simulation details"
\item   Please update Ref. [32] if possible.

Ref. [32] has been accepted in PRB but has not yet been published
\item   FQ: This funding provider could not be uniquely identified during our search of the FundRef registry (or no Contract or Grantnumber was detected). Please check information and amend if incomplete or incorrect.

Je te laisse répondre Gabriel

\end{enumerate}

\end{document}