### 2021-10-08 Talks/group meeting 2021/presentation.tex
weekend

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage{graphicx}
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usepackage[backend=biber]{biblatex}

\addbibresource{Biblio_dbl_quantum.bib}

%\bibliographystyle{stylename}
%\bibliography{Biblio_dbl_quantum}

\title{Group meeting : Cross-relaxation with NV centers ensemble in diamond}
%\author{Clément Pellet-Mary\\ Laboratoire de physique de l'ENS\\ \textit{ENS, Paris}}
\date\today

\mode<presentation> {\usetheme{Rochester}}

\begin{document}
\begin{frame}
\maketitle
\end{frame}
\begin{frame}{The Nitrogen Vacancy Center}
\begin{center}
\includegraphics[scale=.4]{Nitrogen-vacancy_center}
\end{center}

\begin{itemize}
\item 0D fluorescent object with ZPL at 638 nm
\item Controllable and readable spin at room temperature (!)
\item Working with 10$^9$ emitters (typ.)
\end{itemize}
\end{frame}
\end{document}