### 2022-04-05 Manuscrit Thèse/These.tex
toto

\documentclass[a4paper]{article}
\usepackage[french]{babel}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}

\title{Le titre}

\begin{document}
\maketitle
\section{Section 1}
\subsection{Subsection 1}

\end{document}