### 2022-05-06 Manuscrit Thèse/These.tex
tata yoyo

\documentclass[a4paper]{article}
\usepackage[french]{babel}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage[]{braket} % \bra, \ket etc
\usepackage{graphicx}

\title{Le titre}

\begin{document}
\maketitle
\section{Section 1}
\subsection{Subsection 1}

\section*{Appendix}
\subsection{The NV$^-$ center ground sate spin Hamiltonian}
The most general expression of the NV$^-$ ground state spin Hamiltonian can be written as :
\begin{equation}
\mathcal{H}_{NV}= D S_z^2 + \mathcal{H}_{\rm mag} + + \mathcal{H}_{\rm HF} + \mathcal{H}_{\rm elec} + \mathcal{H}_{\rm strain}
\end{equation}
Where : 
\begin{itemize}
\item $D=2.87\ \rm GHz$ is the zero field splitting
\item $\mathcal{H}_{\rm mag}= \gamma_e \vec{B} \cdot \vec{S}$, $\gamma_e=2.8\ \rm MHz/G$
\end{itemize}
\end{document}