### 2022-08-02 Manuscrit Thèse/Glossary.tex
tata

\documentclass[a4paper]{article}
\usepackage[]{amsmath}
\usepackage[]{physics} % \bra, \ket etc
\usepackage{graphicx} %Pour les figures je crois

\begin{document}
\section{Glossary}
\begin{itemize}
\item PL
\item ODMR
\item CR
\item P1/VH-/War1
\end{itemize}
\section{Samples}
\subsection{CVD}
\begin{itemize}
\item CVD-pink : CVD rose d'alex
\item CVD-B/C : cf article de Midrel
\end{itemize}
\subsection{HPHT}
\begin{itemize}
\item SBST-B/C : substrat des CVD-B/C
\end{itemize}
\end{document}