### 2022-08-29 Manuscrit Thèse/chapter 4/Chapter 4.tex
bvjrk

\documentclass[a4paper,11pt]{report}
\usepackage[]{amsmath}
\usepackage[]{physics} % \bra, \ket etc
\usepackage{graphicx} %Pour les figures je crois
\usepackage{hyperref}
\usepackage[
    backend=biber, 
    natbib=true,
    style=numeric-comp,
    sorting=none, %Pour faire apparaitre les refs dans l'ordre
    hyperref=true
]{biblatex} %Imports biblatex package
\addbibresource{Bib_ch4.bib} %Import the bibliography file

\usepackage{amssymb} %quelques symboles dont gtrsim /lesssim
\usepackage{subcaption} % package pour faire des subfigures
\usepackage{multirow} % package pour multirow/multicolumn
\usepackage{booktabs} % package pour top/mid/bottom rule
\usepackage{tcolorbox} % toujours plus de boites
\usepackage{xcolor} % Pour avoir des couleurs dans les équations

\title{}
\begin{document}
\chapter{NV-NV CR under transverse or low fields: application to magnetometry}

(mentionner l'article ici !).... The main motivation for this study was the potential to use NV-NV CR as a low field magnetometry protocol. Indeed, while NV-NV CR lines can be used to perform magnetometry with non zero fields [les russes], there are several advantages to use NV-NV CR as close to the zero magnetic field region, in particular the non dependence of the magnetic field orientation with respect to the crystal lattice. The behavior of

\section{NV spin Hamiltonian under low and transverse fields}

Before looking at the NV-NV CR in the low or transverse field regime, we first need to consider how the general NV physics is modified under those regimes, and in particular we need to look at the modifications of the spin Hamiltonian and the change in the eigenstates.

\subsection{NV spin Hamiltonian in zero external magnetic field}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{Figures/ESR_simus}
\caption{Simulations of inhomogeneous zero field ODMR when sampling various parameters. a) Simulation when sampling each components of the magnetic field over a Gaussian of deviation $\sigma=2\ \rm G$. b) Simulation when sampling each components of the strain tensor $\bar{\bar{\varepsilon}}$ over a Gaussian of deviation $\sigma=2\cdot 10^{-4}$. c) Simulation when sampling each components of the electric field over a Gaussian of deviation $\sigma=2\cdot 10^{5}\ \rm{V/cm}$. d) Experimental ODMR spectrum in zero external field taken on sample ADM-150-2}
\label{simus ESR}
\end{figure}

In the absence of external magnetic field, we have to take into account other elements which would otherwise be of second order in the spin Hamiltonian. These elements are: the random local magnetic fields caused by paramagnetic impurities, the local electric field caused by charged impurities, and the crystal strain \citep{doherty2012theory, udvarhelyi2018spin, mittiga2018imaging}. The hyperfine splitting due to nearby nuclei will be considered separately, although to a large extent it behaves like a local magnetic field.

Due to the large zero field splitting $D=2870 MHz$ between the $\ket{0}$ and $\ket{\pm 1}$ states, we will consider the $\ket{0}$ to always be an eigenstate of the spin Hamiltonian under zero external field (which is equivalent to say that we neglect the terms in $\ket{0}\bra{\pm 1}$ in the spin Hamiltonian). The problem is then reduced to the $\{ \ket{-1}, \ket{+1} \}$ subsystem.

The NV$^-$spin Hamiltonian in the $\{ \ket{-1}, \ket{+1} \}$ basis can be written as \citep{udvarhelyi2018spin}:
\begin{equation}
\mathcal{H}=\begin{pmatrix}
D-\gamma_e B_\parallel + f_\parallel(\mathbf{E}) + g_\parallel(\bar{\bar{\varepsilon}}) & f_\perp(\mathbf{E}) + g_\perp(\bar{\bar{\varepsilon}})\\
f^*_\perp(\mathbf{E}) + g^*_\perp(\bar{\bar{\varepsilon}})&D+\gamma_e B_\parallel + f_\parallel(\mathbf{E}) + g_\parallel(\bar{\bar{\varepsilon}})
\end{pmatrix},
\label{Hamiltonien pm1}
\end{equation}
where $B_\parallel$ is the component of the magnetic field along the NV axis, and $f_\parallel, f_\perp, g_\perp$, and $g_\parallel$ are functions of the electric field $\mathbf{E}$ and the strain tensor $\bar{\bar{\varepsilon}}$, whose expressions are:

\begin{align}
f_\parallel(\mathbf{E})&=d_\parallel E_z, \\
f_\perp(\mathbf{E})&=d_\perp ( E_x + i E_y), \\
g_\parallel(\bar{\bar{\varepsilon}})&= h_{41}(\varepsilon_{xx}+\varepsilon_{yy})+h_{43} \varepsilon_{zz}, \\
g_\perp(\bar{\bar{\varepsilon}}) &= \frac{1}{2} \left[ h_{16}(\varepsilon_{zx}+i \varepsilon_{zy}) + h_{15}\left(\frac{\varepsilon_{yy}-\varepsilon_{xx}}{2}+i\varepsilon_{xy}\right) \right],
\end{align}
where $d_\parallel=0.35 \ \rm{Hz\, cm/V}$ and $d_\perp=17\ \rm{Hz\, cm/V}$ have been measured experimentally \citep{van1990electric}, and $h_{43}=2300\ \rm{MHz}$, $h_{41}=-6420\ \rm{MHz}$, $h_{15}=5700\ \rm{MHz}$ and $h_{16}=19660\ \rm{MHz}$ were computed through DFT \citep{udvarhelyi2018spin} and show reasonable agreement with experiments \citep{barson2017nanomechanical}.

Importantly, as pointed in \citep{mittiga2018imaging} we notice that both the electric field and the strain have a \textit{shifting} component ($f_\parallel$ and $g_\parallel$) which shifts equally both eigenstates of the Hamiltonian, and a \textit{splitting} component ($f_\perp$ and $g_\perp$) which splits in energy the two eigenstates.

The main difference between the electric field and the strain is in the numerical prefactors of these components: for the electric field, the splitting parameter $d_\perp$ is $\sim 50$ times higher than the shifting parameter $d_\parallel$, which will result on average to a strong energy split without much shifting. For the strain however, the splitting parameters $h_{15}$ and $h_{16}$ are only $\sim 3$ times higher than the shifting parameters $h_{43}$ and $h_{41}$. The shift in energy will therefore tend to blur the energy split when averaging over a large number of spins.

Fig. \ref{simus ESR} shows a simulation of how each parameters of the spin Hamiltonian - local magnetic field, local electric field and strain - affects the zero external field ODMR profile. To do these simulations, I sampled each parameters separately $10^6$ times and plotted the histogram of the two eigenvalues of the Hamiltonian written in eq. \ref{Hamiltonien pm1}. Fig. \ref{simus ESR}-d) shows an experimental zero field ODMR spectrum, typical of what we observe with dense NV ensembles.

Experimentally, almost all our samples show the characteristic two bumps in zero external field ODMR. Given the simulation results, the only parameter that can give rise to this shape is the electric field. We will therefore consider that the NV Hamiltonian of our samples is dominated by the local electric field, and more specifically by the transverse electric field $E_\perp\equiv E_x + i E_y$ given the ratio between $d_\perp$ and $d_\parallel$.

We will then adopt the following simplified Hamiltonian for the zero external field regime:
\begin{equation}
\mathcal{H}=\begin{pmatrix}
D&0&d_\perp E_\perp^* \\
0&0&0 \\
d_\perp E_\perp &0&D
\end{pmatrix},
\end{equation}
whose eigenvectors are $\ket{0}$ and$\ket{\pm}$ of eigenvalues 0 and $D\pm d_\perp \abs{E_\perp}$, where $\ket{\pm}$ are defined as:
\begin{equation}
\ket{\pm}=\frac{\ket{+1}\pm e^{-i\phi_E}\ket{-1}}{\sqrt{2}},
\end{equation}
where $\tan(\phi_E)=E_y/E_x$.

\subsection{NV spin Hamiltonian under purely transverse magnetic field}
\label{sec B transverse}
%\begin{figure}[h]
%\centering
%\includegraphics[width=0.7\textwidth]{Figures/map_etats_propres}
%\caption{Closeness of the Hamiltonian most excited state $\ket{3}$ with either $\ket{+1}$ or $\ket{+}$ as a function of the external magnetic field amplitude and angle $\theta$ with respect to the NV axis.}
%\label{map champ transverse}
%\end{figure}

We will consider here the case of purely transverse magnetic field with respect to the NV axis, i.e. $\mathbf{B}=B_x \hat{e}_x + B_y \hat e_y$, and more specifically the regime where $d_\perp E_\perp < \frac{(\gamma_e B_\perp)^2}{D} << D$. In practice, this generally means $20\ \rm G \lesssim B_\perp \lesssim 200\ \rm G$.

In this regime, the NV Hamiltonian eigenstates are similar to the case dominated by the transverse electric field and can be written $\approx \ket{0}, \ket{\pm}$\citep{qiu2021nuclear, qiu2022nanoscale}, of eigenvalues $\approx -\frac{(\gamma_e B_\perp)^2}{D},D$ and $D+\frac{(\gamma_e B_\perp)^2}{D}$,  where:
\begin{equation}
\ket{\pm}=\frac{\ket{+1}\pm e^{-2i\phi_B}\ket{-1}}{\sqrt{2}},
\end{equation}
and $\tan(\phi_B)=B_y/B_x$.

For the case where $d_\perp E_\perp \sim \frac{(\gamma_e B_\perp)^2}{D}$ and $\phi_E \neq 2\phi_B$, the eigenstates of the Hamiltonian are still $\ket{0},\ket{\pm}$ with a relative angle $\phi$ in between $\phi_E$ and $2\phi_B$.

In conclusion, whenever the spin Hamiltonian is dominated by a transverse field, either electric or magnetic, we can consider that the eigenstates of the spin Hamiltonian or $\ket{0}, \ket{-}$ and $\ket{+}$, whereas when the Hamiltonian is dominated by the longitudinal magnetic field, its eigenstates are $\ket{0}, \ket{-1}$ and $\ket{+1}$.

\subsection{Hyperfine coupling and inhomogeneous broadening}
\label{sec modif T2*}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{Figures/Comparaison_ESR_T2}
\caption{tata. Changer les T1 avec T1ph=5ms}
\label{ESR for T2*}
\end{figure}

We will now look at the modification in the ODMR linewidths caused by the different magnetic field regimes. These change are relevant to our study of NV-NV CR due to the relation between the dipole-induced relaxation rate and $T_2^*$ detailed in sec. [REF].

A consequence of the change in the Hamiltonian eigenstates from the $\{ \ket{0}, \ket{\pm 1} \}$ to the $\{ \ket{0}, \ket{\pm} \}$ basis is that the Hamiltonian eigenvalues are sensitive to different part of the environment. In the $\{ \ket{0}, \ket{\pm 1} \}$, the eigen energies are linearly sensitive to (longitudinal) magnetic field, and only sensitive to electric fields at the second order, and vice versa for the $\{ \ket{0}, \ket{\pm} \}$ basis.

These different sensitivities affect both the inhomogeneous broadening, due to the local electric and magnetic noise, and the hyper-fine coupling to the various surrounding nuclei. These effects can drastically modify the ODMR lineshape in zero external magnetic field \citep{jamonneau2016competition} or purely transverse magnetic field \citep{qiu2021nuclear,qiu2022nanoscale}.

We will only consider here the hyper-fine splitting caused by the $^{14}$N nuclei of the nitrogen atom forming the NV center. $^{14}$N represents 99.6 \% of natural abundance nitrogen atoms, and it has an $I=1$ nuclear spin. The full $3\times 3$ Hamiltonian of the NV center in these conditions can be written:
\begin{equation}
\mathcal{H}=\mathcal{H}_e + \mathcal{H}_n + \mathbf{S} \bar{\bar{A}} \mathbf{I},
\end{equation}

where $\mathbf{S}$ is the electronic spin operator, $\mathbf{I}$ the nuclear spin operator, $\mathcal{H}_e$ the previously described electronic spin Hamiltonian, $\mathcal{H}_n$ the nuclear spin Hamiltonian and $\bar{\bar{A}}$ the hyper fine tensor. $\mathcal{H}_n$ and $\bar{\bar{A}}$ can be written:

\begin{align}
\bar{\bar{A}} &= \begin{pmatrix}
A_{xx} & 0 & 0 \\
0 & A_{yy} & 0 \\
0 & 0 & A_{zz}
\end{pmatrix} \\
\mathcal{H}_n &= \gamma_N \mathbf{I} \cdot \mathbf{B} + Q I_z^2,
\end{align}

where $\gamma_N=0.308\ \rm kHz/G$, $Q=-4.945\ \rm MHz$, $A_{zz}=-2.162\ \rm MHz$ and $A_{xx}=A_{yy}=-2.62\ \rm MHz$ \citep{smeltzer2009robust}.

Fig. \ref{ESR for T2*} shows ODMR spectra for two samples, CVD-pink and Sumi-2, in three magnetic configuration: for a strong longitudinal magnetic field, where the NV eigenbasis is $\{ \ket{0}, \ket{\pm 1} \}$, and for a strong transverse magnetic field or no magnetic field, where the NV eigenbasis is $\{ \ket{0}, \ket{\pm} \}$. A table with the full width at half maximum of each ODMR line is shown in Fig. \ref{ESR for T2*}-d)

While both samples are relatively equivalent in term of NV concentration ([NV]=$3 \sim 5$ ppm), sample Sumi-2, being a Type 1B HPHT sample, contains significantly more impurities besides NV centers, most likely P1 centers. These impurities cause both magnetic (because of paramagnetic impurities) and electric (because of charged impurities) field noise.

Fig. \ref{ESR for T2*}-a) allows us to evaluate the magnetic field noise in both samples, since the $\ket{\pm 1}$ states are not sensitive to weak electric fields. We indeed find that the Sumi-2 sample has much more magnetic noise, to the point where the hyper-fine structure is no longer resolved. The total width of the line however is mostly dominated by the hyper-fine splitting, which result in a similar total linewidth in both cases

We should note that the current practice in magnetometry is to consider only of the three hyper-fine lines when they are resolved. This is because magnetometry protocol usually relies on a microwave field with a very well defined frequency, which can effectively select only one of the lines. In our case however, we have to consider the spectral overlap between NV centers and fluctuators, which have an additional broadening of $2\gamma_f \approx 6\ \rm MHz$ [REF] that completely obscures the hyper-fine structure. We therefore have to consider the full linewidth, even when the hyper-fine structure is resolved.

Fig. \ref{ESR for T2*}-b) allows us to evaluate the electric field noise in both samples, since the $\ket{\pm}$ states are not sensitive to weak magnetic fields. Similarly, the hyper-fine structure is hidden in this configuration since all three hyper-fine levels are nearly-degenerate, provided that $\frac{(\gamma_e B_\perp)^2}{D} > A_{xx},A_{zz},Q$ which is typically the case for $B_\perp > 40\ \rm G$. We can note that the electric field noise is significantly stronger in the HPHT sample, which leads to an ODMR linewidth more than twice as big.

Finally, \ref{ESR for T2*}-c) shows the linewidth of both samples for zero external magnetic field. For sample Sumi-2, the profile and linewidth is similar to the case of the purely transverse magnetic field, which is consistent with the fact that the electric field noise is stronger than the residual magnetic fields (earth magnetic field and hyper-fine interaction). For sample CVD-pink however, the electric field noise is smaller than the hyper-fine interaction, meaning that only the $\ket{m_I=0}$ states (the ones closest to the central dip) will be dominated by the electronic noise, and be in the electronic states $\ket{m_e=\pm}$. The $\ket{m_I=\pm 1}$ states are dominated by the hyper-fine field and are in the electronic $\ket{m_e=\pm 1}$ basis.

Overall, we are mostly interested in the linewidth of the ODMR lines for various magnetic field conditions. While these linewidth are quite significantly changed for sample CVD-pink, the change is less pronounced on sample Sumi-2. Most HPHT samples we used are from type 1B diamond, and behave similarly to sample Sumi-2. Since almost all measurements in this chapter will be conducted on HPHT samples, we will consider the modification of $T_2^*$ as a minor effect.

\section{NV-NV CR in the low magnetic field regime}
In this part, we will study the cross-relaxation between NV centers under low magnetic field, whose spin Hamiltonian is dominated by a transverse electric field. Understanding the various mechanisms operating in this regime will be important for the magnetometry protocol presented after.

\subsection{Experimental observations}
\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{Figures/scan_1x1x1x1}
\caption{tata. Changer les T1 avec T1ph=5ms}
\label{scan 1x1x1x1}
\end{figure}
We will start by showing that NV-NV CR behaves differently in the low magnetic field regime compared to the longitudinal field dominated regime which we studied in the last chapter.

The main issue with studying NV-NV CR in low to zero magnetic field is that there are many competing effects happening simultaneously, with few buttons to adjust to isolate each effects.

Fig. \ref{scan 1x1x1x1}-c) and d) show the evolution of the NV PL and stretched lifetime $T_1^{\rm dd}$, defined in the last chapter [REF], as the magnetic field is scanned from 0 to 60 G. The ODMR at the initial and final magnetic fields are shown in Fig. \ref{scan 1x1x1x1}-a) and b).

While it is clear that the spin lifetime as well as the PL increases with the magnetic field, there is no clear indication that this is because of the specificity of the low field region. Indeed, the most likely explanation in this case is that the four classes of NV centers get split apart as the magnetic field increases, which reduces the density of resonant fluctuators for each NV centers.

We can note however that the PL is not an exact mirror of the spin dynamics: while the low field drop in PL is indeed associated with a reduced spin lifetime, the drop in PL at higher magnetic field comes from the states mixing induced by the transverse magnetic field, and is not associated with a modification of the spin dynamics (to the first order).

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{Figures/scan_100}
\caption{Same measurements as Fig. \ref{scan 1x1x1x1}, still on sample ADM-150-1, but with $\mathbf{B}$ along the [100] axis. Changer les T1 avec T1ph=5ms}
\label{scan 100}
\end{figure}

Fig. \ref{scan 100} presents a way to circumvent this issue: by applying the magnetic field along the [100] crystalline axis, we can make sure that the four classes of NV centers always stay resonant regardless of the magnetic field amplitude.

We can notice that there still is a decrease of both the PL and $T_1^{\rm dd}$ in low field, although considerably smaller than the previous case: in Fig. \ref{scan 1x1x1x1}, $T_1^{\rm dd}$ was reduced by a factor of [REF] in zero field, whereas in Fig. \ref{scan 100}, it was only reduced by a factor of [REF]. The main reason for the PL and $T_1$ drop was indeed the co-resonance between the four classes.

Nevertheless, the fact that there is a drop in zero field when $\mathbf{B} \parallel [100]$ cannot be explained by considering only the inter-class resonances. There are some additional depolarization mechanisms which are proper to the zero field region. We should also note that, while the zero-field PL contrast is bigger in Fig. \ref{scan 1x1x1x1}-c) than in Fig. \ref{scan 100}-c), the PL slope, which is the limiting factor for sensing ability, is actually very similar in both cases with a value $\sim [REF]\ \rm G^{-1}$.

We can also note a drop in PL and a corresponding increase of $1/T_1^{\rm dd}$ for $B \sim 20\ \rm G$ which corresponds to the NV$-^{13}$C-NV cross-relaxation discussed in [REF]

\subsection{Potential causes for low field depolarization}

We will discuss here three possible reasons for the zero field depolarization observed in Fig. \ref{scan 100}. Once presented, we will try to hierarchized the contribution of each of these effects.

\subsubsection{Eigenstates modification}
The first explanation is the modification of the dipole-dipole interaction caused by the change of the NV Hamiltonian eigenbasis from $\{ \ket{0}, \ket{\pm 1} \}$ when $\mathbf{B} \neq 0$ to $\{ \ket{0}, \ket{\pm} \}$ when $\mathbf{B} = 0$.

This modification arise from the new form of the dipole-dipole Hamiltonian in the $\{ \ket{0}, \ket{+}, \ket{-} \} \times \{ \ket{0}, \ket{+}, \ket{-} \}$ basis. We justify this change of basis, where we only considered the single NV Hamiltonian instead of the full two-spins Hamiltonian, by the fact that we are in the weak coupling regime, where $\expval{\mathcal{H}_{dd}} \approx 50\ \rm kHz \ll \frac{1}{2\pi T_2^*} \approx 5\ \rm MHz$ meaning that we can treat the dipole-dipole interaction perturbatively. To compute the decay rate with these new eigenstates, we now need to consider the $\mel{0,\pm}{\mathcal{H}_{dd}}{\pm ,0}$ matrix elements instead of the $\mel{0,\pm 1}{\mathcal{H}_{dd}}{\pm 1,0}$ ones.

The computation of the decay rates in the new eigenbasis involve the averaging of these matrix elements which is detailed in appendix [REF]. The computation in this case is complicated by the fact that the transverse field (either $\mathbf{E}$ or $\mathbf{B}$) responsible for the splitting of the $\ket{\pm}$ levels breaks the Hamiltonian symmetry in the $(xy)$ plane. This means that the dipole-dipole coupling between two spins will depend on their relative $x$ and $y$ axis, defined by the local transverse field, on top of the relative $z$ axis defined by the NV axis.

We therefore need to make an assumption on the distribution of the transverse field in the sample. In zero external magnetic field where the dominant transverse field comes from randomly spaced charged impurities, we can expect the $x$ and $y$ axes to be randomly sampled in their respective $(xy)$ plane. However, since the NV-fluctuator CR is dominated by the closest neighbor of each spin (due to the $1/r^6$ scaling in eq. [REF]), there could still be local correlations in the transverse field felt by the NV and its nearest fluctuator.

We then computed the decay rates for the two extreme cases: first, we consider that the $x$ and $y$ axis of each spin is randomly sampled, which correspond to a correlation length of the transverse field $l_c=0$, and secondly we considered the case where the dominant transverse field as a fixed orientation in the whole sample corresponding to $l_c=\infty$.

We found for $\mathbf{B}=0$ that the expected decay rate was $\Gamma_1=51.4 \Gamma_0^{\rm th}$ if $l_c=0$ and $\Gamma_1=55.0 \Gamma_0^{\rm th}$ if $l_c=\infty$. $\Gamma_0^{\rm th}$ has the same definition as in table [REF], it is the expected decay rate for an isolated class in the $\{ \ket{0}, \ket{\pm 1} \}$ basis.

In both cases, this is a moderate ($\sim 20 \%$) increase compared to $\Gamma_1=42.8 \Gamma_0^{\rm th}$ which we previously found for $\mathbf{B} \parallel [100]$, where all four classes were degenerate but the Hamiltonian eignebasis was $\{ \ket{0}, \ket{\pm 1} \}$. The change in the Hamiltonian eigenbasis for low field is therefore a possible candidate to explain the low field depolarization in Fig. \ref{scan 100}

\subsubsection{Double flips}

\begin{table}[htbp]
\centering
\caption{Simulated depolarization rate for flip-flops and double flips in zero magnetic field}
 \label{table double flips}
\begin{tabular}{c|cc}
%\toprule
{} & $l_c=0$ & $l_c=\infty$ \\
\midrule
$\bar{G}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
$\bar{G}^2=0.5$ & $NA \Gamma_0^{\rm th}$ & $NA \Gamma_0^{\rm th}$ \\
$\bar{G}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
%\bottomrule
\end{tabular}
\end{table}

Another effect which happens only at low magnetic field is the possibility of double flip between the NV center and the fluctuator, which is the simultaneous flip of both spins in the same directions, for example going from $\ket{0,-1}$ to $\ket{+1,0}$. The processes $\ket{0, \pm 1}\bra{\mp 1,0}$ are resonant for $B=0$ and not anymore when the degeneracy between the $\ket{\pm 1}$ states is lifted by the magnetic field.

We have seen however that for weak magnetic fields the eigenstates of each NV Hamiltonian are $\ket{\pm}$ and not $\ket{\pm 1}$, due to the transverse electric field. And the $\ket{+}$ and $\ket{-}$ states are not resonant in zero magnetic field. But we can see from Fig. \ref{simus ESR}-d) or \ref{ESR for T2*}-c) that the splitting $2 d_\perp E_\perp \approx 8\ \rm MHz$ is of the same order than the $1/T_1^{\rm dd}$ profile of linewidth $\Gamma^{\rm dd} \approx 8.8\ \rm MHz$ measured on Fig. [REF]. In zero external magnetic field, the $\ket{+}$ and $\ket{-}$ states are therefore close enough in energy that double flip can take place.

To compute the depolarization induced by the double flips, we need to add another decay channel in the fluctuator model, on top of the flip-flop one. We also need to take into account the fact that the $\ket{+}$ and $\ket{-}$ states are not fully resonant, which will modify the $\bar{G}$ factor defined in [REF]. Finally, the question about the correlation  length of the electric field also needs to be asked.

The results are summed up in Table \ref{table double flips}. We again give the results for the two extreme cases regarding the correlations in the electric field, and we show three possible scenario regarding the pseudo-resonance of the $\ket{+}$ and $\ket{-}$ states, represented by the $\bar{G}^2$ factor: $\bar{G}^2=1$ means that the states are fully resonant ($\omega_+-\omega_- \ll \Gamma^{\rm dd} $), this would be the maximum possible decay rate involving the double flips. $\bar{G}^2=0.5$ represents the case where the splitting between the states is equal to the fluctuator linewidth ($\omega_+-\omega_- \approx \Gamma^{\rm dd}$). This is close to the the experimental values we have. Finally $\bar{G}^2=0$ correspond to the case where the states are too far detuned for the double flips ($\omega_+-\omega_- \gg \Gamma^{\rm dd} $). We come back to the results of the last section where only considered the flip-flops.

We therefore predict that the double flips amounts for an increase of the decay rate by [REF] \% (taking $\bar{G}^2=0.5$) compared to the previous case.

\subsubsection{$T_2^*$ modification}

The last aspect we considered was the change in $T_2^*$ (we include the change in the hyper-fine interaction in the $T_2^*$) for weak magnetic field, as discussed in sec. \ref{sec modif T2*}.

To quantify its impact in the depolarization rate, we employ eq. [REF], wherche use the values $2\gamma_f = 6.5\ \rm MHz$ and we take for $\Gamma_f=\Gamma_NV$ the values of half width at half maximum found in Fig. \ref{ESR for T2*}.

We then find that the modification of $T_2^*$ should increase the decay rate in zero field by $\sim 6 \%$ for sample Sumi-2, and by $\sim 19 \%$ for sample CVD-pink.

\subsubsection{Summary}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{Figures/Shema_summary_theory}
\caption{Visual summary of the different depolarization effects in low magnetic field and their predicted depolarization rate}
\label{summary_theory}
\end{figure}

Fig. \ref{summary_theory} recapitulate the various mechanisms at play in the low field depolarization of NV ensemble.

We start from from the simplest case where every classes are sepctrally isolated and the only dipole-induced relaxation coms from the flip-flop between NV and fluctuators from the same class. We then consider the case where all four classes are resonant ($B\parallel [100]$) and finally the case B=0.

The three extra depolarization mechanisms in zero field are then detailed, with the predicted increase of depolarization rate for each step. The predicted rates in zero field depend on several parameters, we chose the following ones to compute the different numerical values: $l_c=0$, $\bar{G}^2=0.5$, $2\gamma_f = 6.5\ \rm MHz$ and $2/T_2^*=5.9\ \rm MHz$ (corresponding to sample Sumi-2 if we assume a Lorentzian ODMR profile).

\subsection{Other potential causes}
We have also considered two other causes that we have ultimately ruled out in this case.

\subsubsection{Magnetic field alignment}
\begin{figure}[h]
\centering
\includegraphics[width=.7\textwidth]{Figures/alignement_rose}
\caption{PL of sample CVD-pink as a function of the magnetic field for various misalignment of the field direction with the [100] axis}
\label{alignement rose}
\end{figure}

The first is a potential misalignment of the magnetic field with the [100] axis, which would cause a splitting of the 4 classes as the magnetic field increases. This hypothesis is not consistent with the observations in Fig. \ref{scan 100}-d): we would expect a small  misalignment to cause a slow but steady decrease of $1/T_1^{\rm dd}$ as the magnetic field is increased. Instead we observe a sharp decrease followed by a pseudo-plateau.

We still tried to estimate the sensitivity of the low field depolarization with the field orientation. Fig. \ref{alignement rose} shows PL scans on sample CVD-pink for various alignment of the magnetic field around the [100] axis. We estimate our alignments to be precise within $\pm 1^\circ$. We can see that the central feature is almost unaffected by the orientation within [$-3^\circ$:$+3^\circ$] which confirms that the low field depolarization does not come from a field misalignment.

\subsubsection{Laser polarization}
\begin{figure}[h]
\centering
\includegraphics[width=.7\textwidth]{Figures/pola_laser}
\caption{tata}
\label{pola laser}
\end{figure}

Other studies \citep{anishchik2015low, filimonenko2020weak} previously saw a correlation between the zero magnetic field PL dip and the polarization angle of the laser with respect to the magnetic field.

Fig. \ref{pola laser} shows PL scans on sample ADM-150-3 for various polarization angle of the incident laser. We chose to apply the magnetic field in the laser polarization plane, in a direction that did not match any particular crystalline planes.

We can see no clear differences between the different polarization angles used, whereas \citep{anishchik2015low} and \citep{filimonenko2020weak} saw an antiline growing inside the PL dip when $B \perp E_{\rm las}$. We think that the difference between our experiments come from the fact that we are using denser NV ensembles, and that any effects tied to the laser polarization are hidden by the stronger effects discussed previously.

\section{NV-NV CR under purely transverse magnetic field}
We now want to experimentally discriminate these different contributions and to see how close the actual decay rates are to the predicted ones.

One of the reason we want to separate these contributions is that they scale differently with the different variables of the sample ($T_2^*$, $\gamma_f$ , magnetic and electric noise ...). Understanding which of these effect dominates the zero field dynamics might give us insight on which parameter to optimize if one wants to avoid the zero field depolarization, or on the contrary to increase it.

The ideal way to isolate each contributions would be to apply an electric field strong enough to split the transition of the 4 classes, and to measure the decay rate of a single class dominated by the transverse electric field. Such an electric field would need to be of the order of $\sim 10^6\ \rm{V/cm}$, which is several order of magnitudes greater than the breakdown voltage of air and out of our experimental reach.

Instead of an electric field, we will use a transverse magnetic field, which as described in sec. \ref{sec B transverse} leads to a spin Hamiltonian similar to the one where the transverse electric field dominates, or at least to similar eigenstates.
\subsection{Principle of the experiment}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{Figures/transis_transverse_field}
\caption{Simulation of the eigenstates of the NV Hamiltonian for $d_\perp E_perp=4\ \rm MHz$ and a purely transverse magnetic field $B_\perp$. (a) Eigenvalues for the spin Hamiltonian as a function of $B_\perp$. The three states are labeled $\ket{1}$, $\ket{2}$ and $\ket{3}$ (b) Splitting between the $\ket{2}$ and $\ket{3}$ states (blue curve) and closeness factor $\abs{\bra{3}\ket{+}}^2$ between the $\ket{3}$ and $\ket{+}$ (red curve)}
\label{eigenstates transverse field}
\end{figure}

Fig. \ref{eigenstates transverse field} shows a simulation of the eigenstates and eigenenergies for a NV Hamiltonian subject to transverse electric field and magnetic field. For convenience we chose to take $E_\perp = E_x$ and $B_\perp = B_x$, taking other geometric configuration only alters slightly the results. We labeled the eigenstates of the Hamiltonian $\ket{1}$, $\ket{2}$ and $\ket{3}$ in ascending order of energy.

Fig. \ref{eigenstates transverse field}-b) shows both the splitting $\Delta \nu$ between the $\ket{2}$ and $\ket{3}$ states, and how close the $\ket{3}$ state is to the $\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}}$ state via the factor $\abs{\bra{3}\ket{+}}^2$. We look at the $\ket{3}$ state since $\ket{2}=\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}}$ in this case.

These two metrics, $\Delta \nu$ and $\abs{\bra{3}\ket{+}}^2$ are what we are interested in: we want to increase $\Delta \nu$ to the point where the double flips are completely quenched, but we want $\abs{\bra{3}\ket{+}}^2$ to remain close to 1 since we wanted to observe the modification caused by the eigenbasis $\{\ket{0},\ket{\pm}\}$. This way we can isolate the $T_1^{\rm dd}$ contribution coming from the change of eigenstates to the one coming from the double flips. Based on the plots in Fig.\ref{eigenstates transverse field}-b), the region with $B_\perp \in$ [100,200] G seems to satisfy both these criteria.

\subsection{Experimental data}

\begin{figure}[h!]
\centering
\includegraphics[width=.8\textwidth]{Figures/scan_T1_perp}
\caption{Experimental data for the depolarization under purely transverse magnetic field on sample ADM-150-2. a) ODMR spectrum for $|\mathbf{B}|=20\ \rm G$. The transitions of the class orthogonal to the magnetic field are labeled $\nu_+$ and $\nu_-$. b) Dependency of $\nu_+$ and $\nu_-$ with the magnetic field, as measured through ODMR. c) Measurement of $T_1^{\rm dd}$ as a function of the transverse magnetic field. The corresponding energy splittings $\Delta \nu=\nu_+-\nu_-$ are written on top. We divided the plot between the A region where $1/T_1^{\rm dd}$ decreases, and the B region where it reaches a plateau with a value $1/T_1^{\rm dd}=185 \pm 5\ \rm s^{-1}$ indicated by a red line. We also indicate in green the value found for a purely longitudinal magnetic field $1/T_1^{\rm dd}=126 \pm 10\ \rm s^{-1}$}
\label{champ tranverse exp}
\end{figure}

We perform here a $T_1^{\rm dd}$ measurement similar to the one presented in Fig. [REF], on the same sample and with the same fitting parameters, but this time probing a class orthogonal to the magnetic field.

Fig. \ref{champ tranverse exp}-a) and b) show the evolution of the transition frequencies $\nu_+=\frac{E_{\ket{3}}-E_{\ket{1}}}{h}$ and $\nu_-=\frac{E_{\ket{2}}-E_{\ket{1}}}{h}$ where $\ket{1}$, $\ket{2}$ and $\ket{3}$ are the states described in Fig. \ref{eigenstates transverse field}. The magnetic field was scanned between 20 and 130 G by increment of $\sim 0.5\ \rm G$.

Fig. \ref{champ tranverse exp}-c) shows the evolution of $1/T_1^{\rm dd}$ with the magnetic field. We have denoted two region: in region A, $1/T_1^{\rm dd}$ decreases with the magnetic field, from a value of $350\ \rm s^{-1}$ to a value of $185\ \rm s^{-1}$. In region B, the value of $1/T_1^{\rm dd}$ is stable and does not decrease further when $B_\perp$ or $\Delta \nu=\nu_+=\nu_-$ are increased.

\subsubsection{Region A}

We attribute the decrease in region A to the double flips: as the magnetic field and $\Delta \nu$ are increased, the double flips become less and less resonant to the point where they become completely quenched for $B_\perp \approx 90\ \rm G$ or $\Delta \nu \approx 30\ \rm MHz$. This value is coherent with the previously measured fluctuator linewidth. We can then estimate the part of the double flips in the low field (20 G here) depolarization rate and find that they amount for $\sim 50\%$ in this case.

Another possibility to explain this drop could be that the class orthogonal to $\mathbf{B}$ still exerts flip-flops with the other classes at low magnetic field, as we can see that they are not very far detuned in Fig. \ref{champ tranverse exp}-a). However we have to consider that the detuning between the two states $\nu_+-\nu_-= 9\ \rm MHz$ is significantly lower, even at $B_\perp=20\ \rm G$, than the detuning with the closest class $\nu_\pm-\nu_{\rm other}= 22\ \rm MHz$; and also that there seem to be a small plateau for $1/T_1^{\rm dd}$ at the beginning of region A which is consistent with the double flip hypothesis since  $\nu_+-\nu_-$ is almost flat in this region while $\nu_\pm-\nu_{\rm other}$ increases linearly with the magnetic field.

\subsubsection{Region B}

We now turn to region B. Even tough the double flips are completely quenched, the effects of the eigenbasis modification and the $T_2^*$ reduction are still present, because we can see on Fig. \ref{eigenstates transverse field}-b) that the eigenstates of the spin Hamiltonian are still very close to $\{ \ket{0}, \ket{\pm} \}$ for these magnetic field values.

To evaluate the contribution of the eigenbasis and $T_2^*$, we measure the same class of NV centers, but this time with a longitudinal magnetic field ($B_\parallel \sim 100 G$) to get a baseline value of $1/T_1^{\rm dd}$ in the $\ket{\pm 1}$ basis. We found a value of $1/T_1^{\rm dd}=126 \pm 10\ \rm s^{-1}$ in this case, reported on Fig. \ref{champ tranverse exp}-c) as a green line.

We do find that the decay rate is higher in the $\ket{\pm}$ basis by $\sim 50\% $ which corroborates our predictions, as both the change of eigenbasis from $\ket{\pm 1}$ to $\ket{\pm}$ and the increase of $T_2^*$ in the $\ket{\pm}$ basis are predicted to increase the depolarization rate. The theoretical increase however overestimates once again the actual increase: for the change of eigenbasis alone we predicted an increase of the decay rate by a factor of 4 (see appendix [REF]), over 8 times more than the actual increase.

We unfortunately cannot experimentally discriminate the depolarization coming from the eigenbasis change or the $T_2^*$ while using the same sample, since both of this properties are tied to the Hamiltonian eigenstates. But we do know that in this scenario, for a single isolated class, the contribution of both the change of eigenbasis and $T_2^*$ amounts for $\sim 3$ times less than the double flips.

\subsubsection{Conclusion of the experimental observation}

We have here quantitatively measured different depolarization contributions for a single class submitted to purely transverse magnetic field. We expect that the depolarization for low magnetic fields behave similarly, and that the difference in NV polarization between the $\mathbf{B}=0$ region and $\mathbf{B} \neq 0$ region is caused by (in order of importance):
\begin{enumerate}
\item The lift of degeneracy between the four classes caused by the projection of the magnetic field on the different NV axis.
\item The double flips between the nearly resonant $\ket{+}$ and $\ket{-}$ states at low magnetic field.
\item The change of the Hamiltonian eigenbasis from $\{\ket{0}, \ket{\pm} \}$ in the electric field dominated region to $\{\ket{0}, \ket{\pm 1} \}$ in the magnetic field dominated region, and the incident change in $T_2^*$.
\end{enumerate}

\section{Introduction to ensemble NV magnetometry}
Before we explore the application of low field depolarization to magnetometry, we must discuss the general field of NV magnetometry.

NV magnetometry is generally decomposed between single NV and ensemble magnetometry. Single NVs,often associated with scanning probes, offer a spatial resolution limited only by how far the magnetic source can be from the NV center (typically $\sim 10\ \rm nm$), but at the cost of a relatively low sensitivity (typically $\sim \mu \rm{T} / \sqrt{\rm Hz}$ \citep{pelliccione2016scanned}).

Ensemble NV centers center on the other hand offer sensitivities below $1\ \rm pT / \sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}, at the cost of working working with samples of several hundreds of $\mu$m.

We will only focus here on NV ensemble magnetometry since we want to give elements of comparisons with our technique based on NV-NV CR. We will not try to give a complete overview of this field, but only to the elements close to the technique we present. For a general overview on microwave-based ensemble NV magnetometry, we recommend the excellent review from Barry et al \citep{barry2020sensitivity}.

\subsection{AC and DC magnetometry}
The most sensitive NV magnetometry protocols use the free precession of the spins to map a small change in the Zeeman energy into the phase of the spins, which is then converted in population. These protocols are generally distinguished between DC-broadband and AC-narrow band.

For DC-broadband magnetometry, the general technique consist of a Ramsey interferometry experiment, presented in chapter 1, which is limited by the spin inhomogeneous coherence time $T_2^* \lesssim 1\ \rm{\mu s}$ for NV ensembles. The bandwidth of the protocol is limited by the repetition rate of the experiment, generally up to $\sim 100\ \rm kHz$. Although less sensitive in theory, continuous-wave (CW) ODMR and pulsed ODMR are also often employed for DC magnetometry. The state of the art in term of DC ensemble magnetometry is a sensitivity of $\sim 20 pT/ \sqrt{\rm Hz}$ \citep{barry2016optical, chatzidrosos2017miniature}.

For AC magnetometry, the basic protocol consist of a spin echo measurement, also presented in chapter 1, which extends the coherence time $T_2^*$ to $T_2 = 10 \sim 100 \rm{\mu s}$, which can be extended even further to $\sim \rm ms$ times with dynamical decoupling \citep{pham2012enhanced}. Due to the longer ``free" precession time, a larger phase can be imprinted on the spin for a similar magnetic field. The AC magnetic field however has to be in phase with the rephasing pulses, which limits the AC sensing to a relatively narrow band. The state of the art in term of AC sensing is $\sim 1\ \rm pT/\sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}.

Recently, \citep{xie2021hybrid} have achieved new records in both DC ($\sim 200\ \rm fT/\sqrt{\rm Hz}$) and AC ($\sim 10\ \rm fT/\sqrt{\rm Hz}$) NV magnetometry by using flux concentrators which are ferromagnets concentrating the magnetic field in a small volume, similarly to an electromagnet.
\subsection{Low field magnetometry}
Magnetometry for low magnetic field ($\gamma B < d_\perp E_\perp$) is challenging since in this regime the NV eigenstates and eigenenergies do not depend on the magnetic field in the first order. This complication is often lifted  by simply using a bias magnetic field, typically from a permanent magnet.

Some systems however need to be kept under low to zero external magnetic fields, for example to study the J-coupling between spins \citep{sutter2012computational}, or the phase of skyrmions \citep{zazvorka2020skyrmion}. To study these systems with NV centers, we need to find new protocols for low field NV magnetometry.

The main technique for low field magnetometry consist in using circularly polarized microwaves, which can only excite the $\ket{0} \to \ket{-1}$ and $\ket{0} \to \ket{+1}$ transitions, even when these states are not the Hamiltonian's eigenstates \citep{mrozek2015circularly, zheng2019zero, lenz2021magnetic, vetter2022zero}. The best sensitivity achieved with this technique was $250\ \rm pT/\sqrt{\rm Hz}$.

Another technique consist in using the $^{13}C-$NV complex we discussed in sec. [REF] which because of the magnetic field offset caused by the $^{13}C$ nucleus are frist order sensitive to the magnetic field even in zero external magnetic field \citep{wang2022zero}. With natural abundance $^{13}C$ however, this means that only $\sim 3\%$ of the NV centers are used, and increasing the $^{13}C$ concentration can lower the coherence of the NV centers. This technique is more suited for a carefully chosen single NV center.

\subsection{Microwave-free magnetometry}
All the precedent magnetemetry protocols described relied on the use of a controllable magnetic field, to coherently manipulate the NV spin state and to measure the action of the magnetic field on the NV Larmor frequency.

There are other protocols which correlate the magnetic field with the spin $T_1$ time (relaxometry). These protocols operate without microwave field and work for either DC or high frequency AC sensing.

\subsubsection{microwave-less DC magnetometry}
The DC microwave-free protocols rely on sharp feature in the PL for specific magnetic fields around certain cross-relaxations processes or level anti-crossing. Fig. [REF] in chapter 2 show such PL features, in particular the ground state level anti-crossing (GSLAC, simply labeled LAC on the figure), the sharpest of these features.

The GSLAC occurs for a longitudinal magnetic field when $\gamma B_z = D$, which corresponds to $B=1024\ \rm G$. In these conditions, the normally bright $\ket{0}$ state and the dark $\ket{-1}$ state become resonant and any residual mixing between these two states (coming from the electric field, the strain or the residual transverse magnetic field) will cause a drop in the NV polarization, and in the PL \citep{broadway2016anticrossing}.

This protocol has been implemented, first only for longitudinal magnetic field measurement \citep{wickenbrock2016microwave} and then for 3D vector measurement \citep{zheng2020microwave} with a sensitivity $\sim 300\ \rm pT/\sqrt{\rm Hz}$.

Another DC microwave-less protocol was proposed based on NV-NV CR \citep{akhmedzhanov2017microwave, akhmedzhanov2019magnetometry}. This technique reconstructs the external magnetic field by scanning an other magnetic field, similar to Fig. [REF] in chapter 3, and by measuring the position of the NV-NV CR lines. By scanning the magnetic field in different directions, one can then recover the initial 3D magnetic field. The author of this paper did not measure the sensitivity of this protocol, but they predict a shot noise limited sensitivity of $24.7\ \rm nT/\sqrt{\rm Hz}$.

\subsubsection{microwave-less AC magnetometry}
The AC microwave-less protocols rely on a direct excitation of the $\ket{0} \to \ket{-1}$ or $\ket{0} \to \ket{+1}$ transitions by the magnetic field probed. The detection bandwidth is adjusted with an external magnetic field to bring the NV transitions in resonance with the field probed, which has been implemented from $\sim 10\ \rm MHz$ to $27\ \rm GHz$ \citep{magaletti2022quantum}.

Sensitivities as low as a few $pT/\sqrt{\rm Hz}$ have been recorded near 2.87 GHz \citep{wang2022picotesla, alsid2022solid}, although both protocols rely on a secondary controllable microwave field (either for an heterodyne detection or to add additional pulses).

\subsection{Orientation-free magnetometry}
Every protocol described so far requires a precise knowledge  of the transition frequency between the $\ket{0}$ and $\ket{\pm 1}$ states versus the magnetic field. This knowledge implies that the direction of each NV axis in a frame of reference is known, because of the strong anisotropy in the NV Hamiltonian.

For a single crystal, these axes can always be calibrated through ODMR. However, for a polycrystalline diamond, diamond powder or even a single diamond with an erratic motion, the knowledge of these axes is not possible. New protocols need to be found if one wants to perform magnetometry with these materials.

To our knowledge, the only orientation-free protocol for DC magnetometry is to use the change in PL caused by the transverse field, which is a property that depends only weakly on the field orientation, especially with an ensemble of NV centers with multiple axes. This protocol however has an extremely low sensitivity and only had marginal use so far\citep{rondin2012nanoscale, tetienne2012magnetic}.

Orientation-free magnetometry can also be performed for noise sensing, provided that that the noise is sufficiently broadband to directly excite the NV transitions regardless of their Larmor frequencies. The noise can be probed either by measuring the spin $T_1$ time \citep{kolkowitz2015probing, andersen2019electron} (which does not require a knowledge of the larmor frequency), or via the PL \citep{finco2021imaging}.

\section{Low field depolarization magnetometry}
We will now present and characterize the low field depolarization magnetometry (LFDM) protocol which rely on the spontaneous depolarization of dense NV ensemble at low magnetic field. This protocol works for DC or slowly varying ($\lesssim 10\ \rm kHz$) magnetic field, is microwave-free, orientation-free, and operates at low magnetic field, although a small additional field is required.

We will first describe its operation and then characterize it and compare it to the previously mentioned NV magnetometry protocols.

\subsection{Principle of the experiment}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/setup_magneto}
\caption{Principle of the LFDM protocol a) Experimental setup. The confocal microscopy setup is similar to the one described in [REF], the lock-in amplifier (LIA) is used here to add a modulation on the magnetic field, instead of the microwave. Both the PL and the LIA signal are monitored. b) example of LIA signal on sample ADM-15-4 when scanning the magnetic field $B_0$ between -150 and 150 G. c) PL signal corresponding to the LIA signal}
\label{setup magneto}
\end{figure}

Fig. \ref{setup magneto} shows the principle of the LFDM protocol. The experimental setup is similar to the one presented in Fig. [REF] with the exception that there are no microwaves, and that the magnetic field generated by the electro-magnet contains both a DC or slowly variable component $B_0$ and an oscillating component $\delta B$ used for the modulation. We typically use a modulation frequency $f_{\rm mod} \approx 1\ \rm kHz$, the same value we use for ODMR measurements. The main voltage $V_0$ responsible for the DC field and the oscillating voltage $\delta V$ responsible for the AC field are summed via a homemade bias tee.

Fig. \ref{setup magneto}-b) and c) show respectively the PL measured directly on the photodiode and the lock-in signal when the DC magnetic field is slowly varied ($f_{\rm scan}<1\ \rm Hz$) from -150 to +150 G. The magnetic field was randomly oriented with respect to the NV axes and did not match any particular crystalline axis.

Both AC (up to $\sim$ 10 kHz) and DC-magnetometry are possible with this setup. to improve the sensitivity however, an AC field needs to be added for the DC sensing, and inversely a DC field needs to be added for the AC sensing. Both of these added fields are typically of a few G for optimal performances, which limits the ultra-low field applications.

Fig. [A FAIRE] shows how the AC and DC sensing work for LFDM. For DC sensing, we want to use an oscillating field with an amplitude that matches the width of the PL dip $|\delta B| \approx \sigma$, in order to maximize the slope of the lock-in signal around $B_0=0$. For AC sensing, we want to use a static field $B_0=\sigma$ corresponding to the region where the lock-in signal is most sensitive to AC field.

\subsection{Characterization of the LFDM protocol}
We will now look at the performances of the LFDM protocol before comparing it to the other magnetometry protocols. The characterization was only done for DC sensing. The samples used for the characterization were 15 $\mu$m micro-diamond bought as is from Adamas nanodiamond (see the sample section [REF] for more information).
\subsubsection{Sensitivity}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/sensi_alternee}
\caption{Sensitivity measurement of the LFDM on sample ADM-15-4. a) Temporal trace of the LIA signal when applying an alternating external magnetic field of amplitude 35 $\mu$T and frequency $f=1\ \rm Hz$. b) Histogram of the DC measurement over 50 s. The data is fitted with Gaussians of width $1.6\ \rm{\mu T}$.}
\label{sensei alternee}
\end{figure}
We characterize the DC sensitivity by looking at the noise floor on the LIA signal when being on the steepest slope, around $B_0=0$. The following parameters were used in this case: the AC field for the modulation had an amplitude $|\tilde{B}_{\rm AC}|\approx 10\ \rm G$ and a frequency $f_{mod} \approx 1\ \rm kHz$, the laser power was $\approx 1\ \rm mW$ and the collected PL power $\approx 1\ \rm{\mu W}$, and the lock-in low-pass filter time constant was set to $\tau=3\ \rm ms$.

Fig. \ref{sensei alternee} shows the sensitivity measurement protocol. First, the applied magnetic field is calibrated with ODMR spectra, then an alternating DC field of known amplitude (here $\Delta B_0=35\ \rm{\mu T}$) is applied on the sample. Fig. \ref{sensei alternee}-a) shows the temporal trace of the LIA signal in response to this alternating DC field. Knowing the amplitude of the alternating field, we can then convert the voltage signal into a magnetic field amplitude, with in this case a ratio $45 \pm 3\ \rm{\mu T /V}$.

Fig. \ref{sensei alternee}-b) shows an histogram of the measured magnetic field amplitude for both the high and low value of the applied magnetic field. The total acquisition time was 50 s with the external field being alternated every second. The histograms are nicely fitted with Gaussian profiles of width $\sigma=1.5\pm 0.1\ \rm{\mu T}$ which allows us to estimate the DC sensitivity of the protocol:
\begin{equation}
\eta = \frac{\sigma}{\sqrt{2 \tau}}= 116 \pm 10\ \rm{nT} / \sqrt{\rm Hz},
\end{equation}

where $\tau=3\ r=\rm ms$ is the low-pass filter time constant.

We estimate that the sensitivity is mostly shot-noise limited because the same LIA voltage noise ($\delta V \approx 35\ \rm mV$) was observed on a magnetic insensitive region (for $B_0 \approx 100\ \rm G$), and is far above the electrical noise measured when no optical signal is sent to the photodiode ($\delta V < 1\ \rm mV$). Laser amplitude fluctuations could also play a role but the noise seem to grow as the square root of the optical power and not linearly with it.

\subsubsection{Angular dependence of the sensitivity}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/sensi_angulaire}
\caption{tata}
\label{angular sensi}
\end{figure}

We will now look at the dependence of the sensitivity with respect to the magnetic field angle. In particular, we are interested in comparing the sensitivity when $\mathbf{B}\parallel [100]$, where there is no lift of degeneracy between the four classes as $\mathbf{B}$ increases; and a case where $\mathbf{B}\nparallel [100]$ and where there is a lift of degeneracy between the four classes. We previously saw in Fig. \ref{scan 1x1x1x1} and \ref{scan 100} that the PL and $T_1$ contrast was much smaller when $\mathbf{B}\parallel [100]$, but we now focus on the sensitivity which is related to the slope of the PL change, and not directly to the contrast.

Fig. \ref{angular sensi}-a) and b) show the LIA signal when the DC field $B_0$ is scanned from $-150$ G to +150 G on the same sample ADM-15-4, but for two different angle of $B_0$. For the blue curve, the field was scanned along the [100] axis, for the orange curve, it was scanned with an angle of $\sim 20\ ^\circ$ compared to the [100] axis

sensi, dépendence sensi angle, variance d'allan
\subsection{Comparison with other magnetometry protocol}
Budker : mesurer vite deuf les pentes
\section{Conclusion and perspectives}
- Improvement to the sensitivity : Scale up (les deux papiers sur le picoTesla mW, avec les deux systèmes), material optimization (15N !, [NV] concentration, )

- Understand the impact of the various factors on the flip-flop CR and DQ CR: T2*, strain, local electric noise ... Mettre ici les deux adamas du 20210927 avec le petit et le gros DQ, et peut etre aussi les DQ et SQ qui s'inversent sur le substrat. Comprendre

- Application for uneven surfaces, polycrystalline heteroepithaxy, low/no microwave environment (dimaond anvil cells, bio sensing). Gradiometry Degen, Arcizet, MRFM

\printbibliography
\end{document}